Management induced changes of soil organic carbon on global croplands

### 2020-11-09 SOCbudgetPaper/appendix.Rmd
update with spell check, references for discussion part, additional analysis on initialization, update of figures

>>>>

## regional SOC stock in GtC from different sources {#append:regcompare}

```{r captionregSOCtable, echo=FALSE}
captionregSOCtable <- "Modelled as well as data based estimation for global SOC stock in GtC for the first 30 cm of soil aggregated over all land area. Note that SoilGrids, GSOC and WISE do not consider changes over time and rely on soil profile data gather over a long period of time, which makes it hard to pinpoint a specific year to these SOC estimations. In this context they will be compared to modelled data (LPJmL4, this study) for the year 2010."
```

```{r regSOCtable, out.width = "100%", fig.cap = captionregSOCtable, echo=FALSE, fig.pos="H"}
# All defaults
knitr::include_graphics("../ResultNotebooks/Output/Images/reg_comparisonfigure.png")
```

## global SOC gap in GtC for different initialization choices {#append:initcompare}

```{r captioninitSOC, echo=FALSE}
captioninitSOC <- "Global SOC gap in GtC for the first 30 cm of soil for different initialization choices."
```

```{r initSOC, out.width = "100%", fig.cap = captioninitSOC, echo=FALSE, fig.pos="H"}
# All defaults
knitr::include_graphics("../ResultNotebooks/Output/Images/scenario_lu.png")
```

## more short comings

- Fertilizer interaction is not included here by accounting for additional N supply that would alter C:N ratio of the carbon inputs. Tier 2 steady-state method is neglecting fertilizer application, however we would have fertilizer amounts at hand to include them, if proper representation of fertilizer within the method would be possible to add.

- Pasture dynamics are neglected and treated as natural vegetation, which might be -- looking on pasture degradation due to overgrazing -- oversimplified for some spots, but is inline with assumption on pasture SOC stocks done before (see Tier 1 IPCC). (Note that also manure excreted to as well as tillage on pastures are neglected within this analysis, since we focus purely on cropland dynamics.)

- Irrigated areas are not crop specific and irrigation is not restricted to growing periods (since it is very complex to calculate average growing periods). Crop specific growing periods might be possible using LPJmL data.

- Flooded rice area are not represented correctly as parametrization does not hold true for flooded conditions.

- Non-net/Gross land use transitions are not tracked in this study.

- Within cropland we do not track croparea transitions, but rather look at statistical distributions of the crop functional types. Due to crop rotations and missing data on crop specific distributions, these transitions would be any way rather uncertain.

- The disaggregation of manure to build-up areas (in the case of extensive monogastrics) is leading to a lot of displaced manure (?) that is cut off

- It is known that there are mismatches between FAO statistics and LUH areas. As far as possibles there were harmonized within this study.

### 2020-11-09 SOCbudgetPaper/discussion.Rmd
update with spell check, references for discussion part, additional analysis on initialization, update of figures

This study provides an analysis of historic SOC stock changes on cropland. We determine the SOC trends on cropland compared to a counterfactual scenario with a world under natural vegetation under identical historical climatic conditions ($SOC_{natveg}$). Our results show, that human activity lead to cummulative SOC emission of around 37 GtC in 2010. Whereas recent modelling estimates of global SOC emissions indicate an on going increase of SOC emissions (Sanderman et al, Pugh et al.), our study indicates that the global SOC gap is slowly closing due to improved management.

According to Sanderman et al. cummulative SOC emissions since beginning of human cropping activities have been at around 37 GtC for the first 30 cm of the soil with half of it attributed to grazing. It was also pointed out that these results might be conservatively low compared to experimental results. Considering the huge uncertainties in modelling SOC on the global scale, our estimate of 37 GtC in 2010 for cropland emissions only still seems consistent with Sanderman et al. estimates.

Furthermore the results of Sanderman et al. modelled historical trends based on agricultural land expansion without considering SOC variations due to different management systems at all. Pugh et al. considered management effects like tillage and residue recycling in a static way, but neither changes over time nor alignment to oberserved historical data like yields-levels or no-tillage areas were taken into account. The study moreover concludes that productivity gains for crops while increasing yield levels by 18%, do not spill over to a substainal decrease in SOC emissions, but only account for less than 1% change in SOC emissions.

Our study for the first time uses a dynanmic management dataset as driver for SOC dynamics. We show that the moderate global cropland expansion of around 11% between 1974 and 2010 and the resulting depletion of SOC stocks in converted cropland has been outweighted by improved agricultural yields and practises. This is challenging Pugh et al. findings of only small effects due to improved pratices.  
Moreover our sensitivity analysis indicades that (1) recycled residue biomass increases driven by higher yields as well as higher higher recycling rates, (2) improved manure recycling (e.g. due to improved animal waste management system) and (3) the adoption of no tillage practises.

Modelling management effects on the global scale comes however with parametric and structural uncertainties. As pointed out by Keel (2017) and Smith (2019), carbon input calculations are highly sensitive to the choice of allometric functions
determining below and above ground residue estimates from harvested quantities. Keel et al.~questioned whether below ground residues might increase with a fixed root:shoot ratio rather than being independend from productivity gains. Moreover the study pointed out that plant breeding shift allometries and might been not reflected in outdated data sources. Following this argumentation SOC results shown in this study might overestimate actual SOC stocks due to an overestimation in residue biomass, which ratio compared to the harvested organ is normally reduced due to breeding. However, looking into management effect evaluation (see below) there is no indication for a systematic overrating of residue biomass, more likely we are still missing carbon inputs to the soil.

@@

To validate our modelled SOC stocks and stock changes under management, we compare our results to default IPCC stock changes factors (cite) which are based on measurement data for croplands (see \@ref(tab:SCFtable)). To allow for comparison, we aggregated our stock change factors to the four IPCC climate zones.

```{r captionSCFtable, echo=FALSE}
captionSCFtable <- "This table shows different estimates of stock change factors $F_{SCF}$, defined as the SOC stock of managed croplands relative to the counterfactual SOC stock of undisturbed natural vegetation. We compare average values for four IPCC climate zone classifications. IPCC 2006 and 2019 default factors (medium input) as well as altered factors for low input of organic matter without any other subsystem consideration are compared to results for 1990 and 2010 of this study (SOC budget).}"
```

@@

Our estimates correspond very well to the default stock-change factors used for the tier 1 estimation of IPCC, 2006. For the tropical regions the assumptions changed notablly from the guidelines in 2006 to the update in 2019, leaving our results too low in comparison with IPCC, 2019. Considering yield gaps in mainly developing regions in the tropics the default assumption of medium input systems, might be an overestimation of actual SOC state. The additional effect of considering low input of biomass in tropical regions can however not explain the full mismatch to IPCC 2019 values but account for at least 5-7% of it.

Looking again on the time trend our study shows the substantial impact of changed management to the development of $F_{SCF}$.

@@

The worlds SOC stock and its changes are highly uncertain (cite), which is visible by the wide range of global SOC stock estimates (see \@ref(tab:SOCtable)).

```{r captionSOCtable, echo=FALSE}
captionSOCtable <- "Modelled as well as data based estimation for global SOC stock in GtC for the first 30 cm of soil aggregated over all land area. Note that SoilGrids, GSOC and WISE do not consider changes over time and rely on soil profile data gather over a long period of time, which makes it hard to pinpoint a specific year to these SOC estimations. In this context they will be compared to modelled data (LPJmL4, this study) for the year 2010."
```

```{r SOCtable, out.width = "8cm", fig.cap = captionSOCtable, echo=FALSE, fig.pos="H"}
# All defaults
knitr::include_graphics("../ResultNotebooks/Output/Images/TableSOC_comparison.png")
```

The global esimates of SOC stock by this study are on the lower end compared to other modelled results or more data driven estimates. Looking on regional results (fig. SX in supplement), our estimates turn out to be in good agreement for most regions with the largest deviations for boreal areas. Considering that the model was parametrized for croplands, these mismatches are not superising since the temperature effects on decomposition are fundamentally different for permaforst soils. To avoid that this bias influences our results, our study focusses exclusively on cropland soils, excluding most of the boreal zone. Morover, when focussing on SOC changes, pristine natural vegetated areas without human land management under the same historic climatic conditions cancel out in the calculation of SOC emissions.

Our estimates for total SOC stocks of the world, as well as our SOC initialization are dominated by the representation of natural lands and pastures, which are however only estimated in a basic manner. For example, we do not have a differentiated parametrization of nitrogen and lignin content of litterfall for woody and grass type biomes. This leaves carbon inputs and decay behaviour for natural land and pastures rather uncertain. The absolute values of stocks and emissions from land-use change therefore have to be used with caution. Especially in less forested areas the natural land representation might be off, due to parameterization assumptions of the natural litterfall.

Nevertheless total SOC stocks are in a reasonable range and results on SOC stock changes in relative or absolute numbers are not altered by pristine vegetaded areas at all. Moreover, the SOC gap from changed management is only affected from the natural vegetation representation in case of land use transitions. Than former natural land SOC stocks are declining to lower levels. However as pointed out before, due to low cropland expansion rates, there SOC declines are outweighted by the general trend of increasing SOC stocks on existing cropland due to management improvements.
We conducted a sensitivity analysis for a wide range of possible parameter combinations for lignin and nitrogen parameterization of natural litterfall, which tend to change the global SOC stock substantially for different choices
(see appendix). It shows that the general trend of decreasing SOC gap is not altered even under very high estimates for natural SOC stocks.

## Important short commings --> move to appendix

Smaller points and shortcommings:

-This study as the IPCC guidelines suggested has limited her focus to the first 30 cm of the soil profile, leaving changes in the subsoil unnoticed. Nevertheless studies (see Don on tillage) have shown, the subsoil to be a game changer in evaluating total SOC losses or gains for no-tillage systems. It has been argued that for intensivly tilled soils, subsoil SOC is increasing due to the import of carbon rich topsoil to deeper soil layers. Following these argumentation SOC stocks in croplands might even be underestimated.

- Fertilizer interaction is not included here by accounting for additional N supply that would alter C:N ratio of the carbon inputs. Tier 2 steady-state method is neglecting fertilizer application, however we would have fertilizer ammounts at hand to include them, if proper representation of fertilizer within the method would be possible to add.

- Pasture dynamics are neglected and treated as natural vegetation, which might be -- looking on pasture degradation due to overgrazing -- oversimplified for some spots, but is inline with assumption on pasture SOC stocks done before (see Tier 1 IPCC). (Note that also manure excreted to pastures is neglected within these analysis, since we focus purely on cropland dynamics.)

- No tillage adaptaion is neglected on cropland due to less common adoption of no tillage and conservation agriculture. Pastures are assumed to be not tilled at all (propably only heavy managed pastures are tilled with some rotation)

- Irrigated areas are not crop specific and irrigation is not restricted to growing periods (since it is very complex to calculate average growing periods). Crop specific growing periods might be possible using LPJmL data.

- flooded rice area are not represented correctly as parameterization does not hold true for flooded conditions.

- Carbon displacement via leaching and erosion is neglected in this study.

- Non-net/Gross land use transitions are not tracked in this study.

- Within cropland we do not track croparea transitions, but rather look at statistical distributions of the crop functional types. Due to crop rotations and missing data on crop specific distributions, these transitions would be any way rather uncertain.

- The disaggregation of manure to uild-up areas (in the case of extensive monograstrics) is leeding to a lot of displaced manure (?) that is cut off

- It is known that there are mismatches between FAO statistics and LUH areas. As far as possibles there were harmonized within this study.

- "The Tier 2 method does not simulate C change but simply calculates an annual C stock change from the
current C stock to the future steady-state soil C stock calculated based on current conditions." Leading to the fact that our total stock results are highly uncertained.

>>>>

This study provides an analysis of historic SOC stock changes on cropland. We determine the SOC trends on cropland compared to a counterfactual scenario with a world under natural vegetation under identical historical climatic conditions ($SOC_{natveg}$). Our results show, that human activity lead to cumulative SOC emission of around 37 GtC in 2010. Whereas recent modelling estimates of global SOC emissions indicate an ongoing increase of SOC emissions ([@pugh_simulated_2015], [@sanderman_soil_2017]), our study indicates that the global SOC gap is slowly closing due to improved management.

According to [@sanderman_soil_2017] cumulative SOC emissions since the beginning of human cropping activities have been at around 37 GtC for the first 30 cm of the soil with half of it attributed to grazing. It was also pointed out that these results might be conservatively low compared to experimental results. Considering the huge uncertainties in modelling SOC on the global scale, our estimate of 37 GtC in 2010 for cropland emissions only still seems consistent with [@sanderman_soil_2017] estimates.

Furthermore, the results of [@sanderman_soil_2017] modelled historical trends based on agricultural land expansion without considering SOC variations due to different management systems at all. [@pugh_simulated_2015] considered management effects like tillage and residue recycling in a static way, but neither changes over time nor alignment to observed historical data like yields-levels or no-tillage areas were taken into account. The study moreover concludes that productivity gains for crops while increasing yield levels by 18%, do not spill over to a substantial decrease in SOC emissions, but only account for less than 1% change in SOC emissions.

Our study for the first time uses a dynamic management dataset as driver for SOC dynamics. We show that the moderate global cropland expansion of around 11% between 1974 and 2010 and the resulting depletion of SOC stocks in converted cropland has been out weighted by improved agricultural yields and practices. This is challenging [@pugh_simulated_2015] findings of only small effects due to improved practices.
Moreover, our sensitivity analysis indicates that (1) recycled residue biomass increases driven by higher yields as well as higher recycling rates, (2) improved manure recycling (e.g. due to improved animal waste management system) and (3) the adoption of no tillage practices.

Modelling management effects on the global scale comes however with parametric and structural uncertainties. As pointed out by [@keel_large_2017] and [@smith_how_2020], carbon input calculations are highly sensitive to the choice of allometric functions determining below and above ground residue estimates from harvested quantities. [@keel_large_2017] questioned whether below ground residues might increase with a fixed root:shoot ratio rather than being independent of productivity gains. Moreover, the study pointed out that plant breeding shift allometries and might be not reflected in outdated data sources. Following this argumentation SOC results shown in this study might overestimate actual SOC stocks due to an overestimation in residue biomass, which ratio compared to the harvested organ is normally reduced due to breeding. However, looking into management effect evaluation (see below) there is no indication for a systematic overrating of residue biomass, more likely we are still missing carbon inputs to the soil.

@@

To validate our modelled SOC stocks and stock changes under management, we compare our results to default IPCC stock changes factors (@ipcc_2006_2006, @ipcc_2019_2019) which are based on measurement data for croplands (see \@ref(tab:SCFtable)). To allow for comparison, we aggregated our stock change factors to the four IPCC climate zones.

```{r captionSCFtable, echo=FALSE}
captionSCFtable <- "This table shows different estimates of stock change factors $F^{SCF}$, defined as the SOC stock of managed croplands relative to the counterfactual SOC stock of undisturbed natural vegetation. We compare average values for four IPCC climate zone classifications. IPCC 2006 and 2019 default factors (medium input) as well as altered factors for low input of organic matter without any other subsystem consideration are compared to results for 1990 and 2010 of this study (SOC budget)."
```

@@

Our estimates correspond very well to the default stock-change factors used for the tier 1 estimation of IPCC, 2006. For the tropical regions the assumptions changed notably from the guidelines in 2006 to the update in 2019, leaving our results too low in comparison with IPCC, 2019. Considering yield gaps in mainly developing regions in the tropics the default assumption of medium input systems, might be an overestimation of actual SOC state. The additional effect of considering low input of biomass in tropical regions can however not explain the full mismatch to IPCC 2019 values but account for at least 5-7% of it.

Looking again on the time trend our study shows the substantial impact of changed management on the development of $F_{SCF}$.

@@

The world's SOC stock and its changes are highly uncertain (cite), which is visible by the wide range of global SOC stock estimates (see \@ref(tab:SOCtable)).

```{r captionSOCtable, echo=FALSE}
captionSOCtable <- "Modelled as well as data based estimation for global SOC stock in GtC for the first 30 cm of soil aggregated over all land area. Note that SoilGrids, GSOC and WISE do not consider changes over time and rely on soil profile data gather over a long period of time, which makes it hard to pinpoint a specific year for these SOC estimations. In this context they will be compared to modelled data (LPJmL4, this study) for the year 2010."
```

```{r SOCtable, out.width = "8cm", fig.cap = captionSOCtable, echo=FALSE, fig.pos="H"}
# All defaults
knitr::include_graphics("../ResultNotebooks/Output/Images/glo_comparisonfigure.png")
```

The global estimates of SOC stock by this study are on the lower end compared to other modelled results or more data driven estimates. Looking on regional results \@ref(append:regcompare), our estimates turn out to be in good agreement for most regions with the largest deviations for boreal areas. Considering that the model was parametrized for croplands, these mismatches are not surprising since the temperature effects on decomposition are fundamentally different for permafrost soils. To avoid that this bias influences our results, our study focuses exclusively on cropland soils, excluding most of the boreal zone. Moreover, when focussing on SOC changes, pristine natural vegetated areas without human land management under the same historic climatic conditions cancel out in the calculation of SOC emissions.

Our estimates for total SOC stocks of the world, as well as our SOC initialization are dominated by the representation of natural lands and pastures, which are however only estimated in a basic manner. For example, we do not have a differentiated parametrization of nitrogen and lignin content of litterfall for woody and grass type biomes. This leaves carbon inputs and decay behavior for natural land and pastures rather uncertain. The absolute values of stocks and emissions from land-use change therefore have to be used with caution. Especially in less forested areas the natural land representation might be off, due to parametrization assumptions of the natural litterfall.

Nevertheless, total SOC stocks are in a reasonable range and results on SOC stock changes in relative or absolute numbers are not altered by pristine vegetated areas at all. Moreover, the SOC gap from changed management is only affected from the natural vegetation representation in case of land use transitions. Then former natural land SOC stocks are declining to lower levels. However, as pointed out before, due to low cropland expansion rates, their SOC declines are outweighed by the general trend of increasing SOC stocks on existing cropland due to management improvements.
We conducted a sensitivity analysis for a wide range of possible parameter combinations for lignin and nitrogen parametrization of natural litterfall, which tend to change the global SOC stock substantially for different choices
(see appendix). It shows that the general trend of decreasing SOC gap is not altered even under very high estimates for natural SOC stocks.

## Important shortcomings --> might be added as well

The **initialization of SOC stocks** in 1960 is assumed to be in steady state considering the land use pattern and management effects of the initialization year followed by a run-in period of 15 years, till the analysis of the results start. That is inline with IPCC guidelines calling for a run-in period of 5-20 years. However, the results are heavily influenced by it as \@ref(append:initcompare) suggested. Not only the level of the SOC gap is shifted, also the declining trend is much smaller or even not visible at all anymore. For the artificial assumption of starting with potential natural vegetation in 1961 the SOC gap is fluctuating around 10 GtC for the period 1975--2010. For the more moderate assumption, that half of the transition towards the steady-state under land use has already appeared in 1960, the SOC gap is around 23 GtC in 2010. The relative declining trend for the SOC gap between 1975 and 2010 (-12%) is however almost the same as for our historic estimation (-14%). According to LUH [@hurtt_harmonization_2020] cropland area has increased over 62% since 1900 leaving our choice of initialization values too low, since reaching a new steady-state can take decades or even centuries. Having no yield or management data available before our initialization year leaves us with imperfect initialization knowledge, where our initialization choice is probably closer to real world values than starting from a natural vegetated steady-state.

Following the IPCC guidelines this study has limited its focus to the **first 30 cm of the soil profile**. In this regard several aspects are over simplified within our approach. Firstly, distribution of carbon inputs into different soil layers are neglected and all carbon inputs are allocated to the topsoil. This particularly overestimates SOC stocks in the first 30 cm of soil below deeper rooting vegetation, which is certainly the case for most of the woody natural vegetated areas. Consequently, our SOC loss estimates are too high.
Secondly changes to the subsoil due to tillage are neglected, which might be important to noticed, hence studies (see Don on tillage) have shown, the subsoil to be a game changer in evaluating total SOC losses or gains for no-tillage systems. It has been argued that for intensively tilled soils, subsoil SOC is increasing due to the import of carbon rich topsoil to deeper soil layers. Following this arguments SOC stocks in croplands might even be underestimated.

Carbon displacement via **leaching and erosion** is neglected in this study. As pointed out by several studies (cite), the final fate of leached or eroded carbon is uncertain. Whereas for soil quality analysis SOC displacement might play an important role, in this budget approach focusing especially on SOC emissions, displaced but not emitted SOC can be treated as SOC stayed on croplands.

### 2020-11-09 SOCbudgetPaper/introduction.Rmd
update with spell check, references for discussion part, additional analysis on initialization, update of figures

Soil Organic Carbon (SOC), the amount of organic carbon stored through the earth’s soil, is the largest terrestrial carbon pool, exceeding the carbon in the atmospheric and vegetation pools multiple times (Batjes). As such, even small changes in drivers of SOC may thus lead to substantial shifts in earth carbon cycle and influence the amount of CO2 in the atmosphere (ref. permafrost melting). The specific amount of carbon stored in the soil is, however, uncertain, with estimates ranging from 1500 to 2400 GtC for the first meter of the soil profile (Batjes, 1996; more). The quality of SOC maps has markedly improved in recent decades, along with better understanding of the factors driving the global magnitude, distribution, and dynamics of SOC pools. Natural properties like climatic, biophysical, and landscape characteristics clearly play the most important roles to determine SOC variations over space and time.
Human intervention, including land cover change and land management, has however added a further driver to SOC change, which alters terrestrial carbon pools in much shorter time scales and is likely one of the most dominant driver of SOC changes on managed land today (cite). Recent studies identify the anthropogenic SOC debt for the first meter of the soil profile at around 116 GtC (Sanderman et al.), compared to previous estimates between 60-130 GtC  (Lal, 2006). Other studies have focused more closely on spatially disaggregation of SOC changes via advanced digital soil mapping techniques (S-World; Stoorvogel 2, 2017) or better representation of biogeochemical processes within SOC dynamics (Hararuk, 20XX). While providing a rough estimate of the order of magnitude of change, these studies lack a detailed consideration of land management, especially of agricultural activities.
Field-scale models (cite Daycent, RothC, Ecosse, C-Tool) are able to capture these land management impacts by using detailed information on crop yield levels, fertilizer inputs and various other on-farm measures. However, due to the lack of comprehensive global management data as input to these models, scaling up to the global extent remains a complex challenge. Our study combines an spatially explicit estimate of agricultural management data on the global level with a 3-pool SOC model parametrized for global croplands. This allows us to estimate SOC stock change factors, as well as organic carbon flow dynamics within the agricultural system. We thereby consider change in SOC caused by historical land cover change as well as of different agricultural management practices, including residue recycling, manure amendments, irrigation, and tillage. We provide the first global, spatially explicit SOC loss maps that considers detailed agricultural management.
This paper will first introduce in the methods section the basic concept of SOC dynamics as applied in this study. We continue with a detailed description of the global gridded management data used here, including crop production levels, residue recycling rates, manure amendments, and the adoption of irrigation and tillage practices. Lastly, we shortly refer to the concept of stock change factors as outlined in the Tier 1 approach of the IPCC guidelines. In our results section we focus on the SOC dynamics of global croplands by (1) analysing the spatially explicit distribution and depletion of SOC from 1975 to 2010, (2) comparing impacts of different management effects on SOC emissions and (3) quantify global agricultural carbon flows and stocks to compare the importance of various management aspects. Finally, we discuss and compare our findings – including their implications for SOC model development – and conclude with an outlook on the ability of SOC management to mitigate climate change and contribute to negative greenhouse gas emissions.

>>>>

Soil Organic Carbon (SOC), the amount of organic carbon stored through the earth’s soil, is the largest terrestrial carbon pool, exceeding the carbon in the atmospheric and vegetation pools multiple times (Batjes). As such, even small changes in drivers of SOC may thus lead to substantial shifts in earth carbon cycle and influence the amount of CO2 in the atmosphere (ref. permafrost melting). The specific amount of carbon stored in the soil is, however, uncertain, with estimates ranging from 1500 to 2400 GtC for the first meter of the soil profile (Batjes, 1996; more). The quality of SOC maps has markedly improved in recent decades, along with better understanding of the factors driving the global magnitude, distribution, and dynamics of SOC pools. Natural properties like climatic, biophysical, and landscape characteristics clearly play the most important roles to determine SOC variations over space and time.
Human intervention, including land cover change and land management, has however added a further driver to SOC change, which alters terrestrial carbon pools over much shorter time scales and is likely one of the most dominant driver of SOC changes on managed land today (cite). Recent studies identify the anthropogenic SOC debt for the first meter of the soil profile at around 116 GtC (Sanderman et al.), compared to previous estimates between 60-130 GtC  (Lal, 2006). Other studies have focused more closely on spatially disaggregation of SOC changes via advanced digital soil mapping techniques (S-World; Stoorvogel 2, 2017) or better representation of biogeochemical processes within SOC dynamics (Hararuk, 20XX). While providing a rough estimate of the order of magnitude of change, these studies lack a detailed consideration of land management, especially of agricultural activities.
Field-scale models (cite Daycent, RothC, Ecosse, C-Tool) are able to capture these land management impacts by using detailed information on crop yield levels, fertilizer inputs and various other on-farm measures. However, due to the lack of comprehensive global management data as input to these models, scaling up to the global extent remains a complex challenge. Our study combines a spatially explicit estimate of agricultural management data on the global level with a 3-pool SOC model parametrized for global croplands. This allows us to estimate SOC stock change factors, as well as organic carbon flow dynamics within the agricultural system. We thereby consider change in SOC caused by historical land cover change as well as of different agricultural management practices, including residue recycling, manure amendments, irrigation, and tillage. We provide the first global, spatially explicit SOC loss maps that considers detailed agricultural management.
This paper will first introduce in the methods section the basic concept of SOC dynamics as applied in this study. Additionally, we shortly refer to the concept of stock change factors as outlined in the Tier 1 approach of the IPCC guidelines. Lastly, we continue with a detailed description of the global gridded management data used here, including crop production levels, residue recycling rates, manure amendments, and the adoption of irrigation and tillage practices.  In our results section we focus on the SOC dynamics of global croplands by (1) analysing the spatially explicit distribution and depletion of SOC from 1975 to 2010, (2) comparing impacts of different management effects on SOC emissions and (3) quantify global agricultural carbon flows and stocks to compare the importance of various management aspects. Finally, we discuss and compare our findings -- including their implications for SOC model development -- and conclude with an outlook on the ability of SOC management to mitigate climate change and contribute to negative greenhouse gas emissions.
